### 2021-08-21 doc/thesis/chapters/implementation.tex
more of impl section

The transmitter is implemented to provide a reproducable source of data, which can be used to feed the receiver.
Another option would be to use external hardware as a software-defined radio and record some antenna data.
However, the main advantage of a locally implemented transmitter over a recording is, that the raw data input is known.
Thus, the expected output of the receiver model is this raw data, that was previously modulated by the local transmitter.
In the case of a recording, the original data is not known, which does not allow to actually verify the receiver.\\

@@

The next section in the block diagram represents multiple processes in parallel.
The LR Diff part is shifted to 38~kHz in the spectrum.
This is done by modulating the signal with the 38~kHz sub-carrier.
A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
It band-limits the spectrum of the LR~Diff part between 23 and 53~kHz.
The BP filter introduces a constant group delay of the signal.
In order to maintain the alignment of all signals, this side-effect needs to be accounted for and be compensated in the other signal branches.
Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filter's group delay.
The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.\\

@@

The receiver is implemented, following the block diagram in Figure~\ref{fig:bd_dsp_detailed}.
Further assumptions and simplifications, such as the fixed point arithmetic are described into detail above.

@@

The Matlab implementation also includes a comprehensive range of analysis, especially in terms of visual plots.
These plots are useful during the development of new DSP algorithms, in order to verify the correct thereof.\\

An example of a plot that is created by the analysis is depicted in Figure~\ref{fig:matlab_analysis_time_domain}.
The diagram shows the respective signals in the time domain.
Above two timelines display the input data, as it is read from the input audio file.
The two diagrams in the center plot the Mono and LR Diff signals.
The bottom diagrams then represent the result after the entire signal processing chain.
It can be seen that the DSP chain is not able to demodulate the original signal without any error.
However, the result shows a clear separation of the left and right channels, as resembles the original input very closely, which is good enough for this case study.\\

\includepicture [1.0] [0] {Matlab analysis in the time domain.} {matlab_analysis_time_domain} {img/matlab/matlab_analysis_time_domain}

Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
An example thereof is shown in Figure~\ref{fig:matlab_analysis_freq_domain}.
The spectrum of an FM channel multiplex signal, as it is generated by a transmitter before the modulation, is shown in this diagram.
The spectrum after the demodulation through a receiver is shown in the lower diagram.
Spectral parts, such as the pilot tone at 19~kHz, or the LR Diff part around 38~kHz can be seen clearly.
The demodulated spectrum shows an artifact at 57~kHz, which does not exist in the original transmitter spectrum.
This is an artifact of the demodulation process, specifically the frequency-shift of the LR Diff part to baseband.
The artifact is very low in energy, because it is attenuated by a BP filter in the down-conversion process, and can thus be ignored for further processing, even though it can still be seen in the spectral analysis.

\includepicture [1.0] [0] {Matlab analysis in the frequency domain.} {matlab_analysis_freq_domain} {img/matlab/matlab_analysis_freq_domain}

@@

This section describes the implementation in GNU Radio.
A transmitter, as well as a receiver are implemented using this software.
Both can be deployed on actual hardware, so a continuous transmitter or receiver is built in a real environment.

@@

GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
Multiple blocks are connected to a block diagram, that is very similar to the one in Figure~\ref{fig:bd_matlab_transmitter}.
The USRP device is used to broadcast the RF signal over the air.\\

@@

Special care needs to be taken when transmitting signals over the air.
The frequency spectrum is strongly limited by the radio regulations and may be monitored by the authorities.
Please make sure to adapt your setup to the respective local regulations, to avoid any legal issues.
Some of the regulations are mentioned in Section~\ref{sec:FrequencyBand}, but may differ in various countries, however.\\

@@

\includepicture [0.9] [0] [H] {Block diagram of the GNU Radio transmitter setup.} {bd_gnuradio_setup_transmitter} {img/draw.io/bd_gnuradio_setup_transmitter}

@@

The receiver is also implemented, by using standard DSP blocks that are available in GNU Radio.
The implemented structure however does not represent a stereo receiver.
Instead, a mono receiver is implemented.
The antenna signal is FM-demodulated, decimated and then directly sent to the speakers as the output, without any further processing.
This simple receiver was intended as a proof-of-concept and a first example of the cooperation of the external hardware and the GNU Radio software.
Thus, the mono receiver was sufficient, as the focus of the GNU Radio application is put on the transmitter side.\\

This receiver works with both devices, the USRP b200mini and the RTL-SDR.
Figure~\ref{fig:bd_gnuradio_setup_receiver} shows a block diagram of the receiver application.

\includepicture [0.9] [0] [H] {Block diagram of the GNU Radio receiver setup.} {bd_gnuradio_setup_receiver} {img/draw.io/bd_gnuradio_setup_receiver}

@@

The usage of a software like GNU Radio in combination with external hardware is an interesting tool for a use-case like this project.
An FM receiver is to be implemented in multiple different versions.
All of them need to be verified somehow -- this is where the SDR transmitter created with GNU Radio and external hardware opens up an entire new possibility from the viewpoint of system level design integration.\\

Previously it was explained that it is not possible to use recorded antenna data for verification.
The reason was, that the underlying original data is not known and therefor a comparison of the original versus the received, decoded data is not feasible.
Consequently, the Matlab transmitter model was justified by the fact, that its originally encoded audio data is well-known.\\

However, with the current approach of using a GNU Radio SDR transmitter, the encoded data \textit{is} well-known, because it is generated locally.
Thus, the SDR can be used to transmit a verification signal.
This signal could be recorded to file by another SDR, or sourced into the DUT (device under test) directly.

>>>>

The transmitter is implemented to provide a reproducable source of data, which can be used to feed the receiver.
Another option would be to use external hardware as a software-defined radio (SDR) and record some antenna data.
However, the main advantage of a locally implemented transmitter over a recording is, that the raw data input is known.
Thus, the expected output of the receiver model is this raw data, that was previously modulated by the local transmitter.
In the case of a recording, the original data is not known, which does not allow to actually verify the receiver.\\

@@

The next section in the block diagram represents multiple processes in parallel.
The LR Diff part is shifted to 38~kHz in the spectrum.
This is done by modulating the signal with the 38~kHz subcarrier.
A subsequent BP filter serves to remove the artifacts of the modulation, such as the image replica.
It band-limits the spectrum of the LR~Diff part between 23 and 53~kHz.
The BP filter introduces a constant group delay of the signal.
In order to maintain the alignment of all signals, this side-effect needs to be accounted for and be compensated in the other signal branches.
Thus, the Mono signal, as well as the Pilot Tone need to be delayed by the amount of samples defined by the BP filter's group delay.
The subsequent amplifier blocks create the defined proportions of the sub-band, such as 90\% for the audio parts, and 10\% for the Pilot Tone \cite{FmMultiplexingForStereo}.\\

@@

The Matlab receiver is implemented according to the block diagram in Figure~\ref{fig:bd_dsp_detailed}.
Further assumptions and simplifications, such as the fixed point arithmetic and the filter design are already described into detail in the previous chapters.

@@

The Matlab implementation also includes a comprehensive range of analysis functionalities, especially in terms of visual plots.
These plots are useful during the development of new DSP algorithms, in order to verify the correct functionality thereof.\\

An example of a plot that is created by the analysis is depicted in Figure~\ref{fig:matlab_analysis_time_domain}.
The diagram shows the respective signals in the time domain.
Above two timelines display the input data, as it is read from the input audio file.
The two diagrams in the center plot the Mono and LR Diff signals.
The bottom diagrams then represent the result after the entire signal processing chain.
It can be seen that the DSP chain is not able to demodulate the signal without any error, as the output signal does not exactly match the original input signal.
However, the result shows a clear separation of the left and right channels, and resembles the original input very closely, which is good enough for this case study.\\

\includepicture [1.0] [0] {Matlab analysis in the time domain. The stereo audio signal of the voice saying \textit{Left channel, Right channel} can be recognized in the signal, which allows an audible and visual verification of the receiver's functionality.} {matlab_analysis_time_domain} {img/matlab/matlab_analysis_time_domain}
%TODO: latex font for this diagram!

Further analysis is done in the frequency domain, especially in order to better understand the modulation of the different spectral parts in the FM channel.
An example thereof is shown in Figure~\ref{fig:matlab_analysis_freq_domain}.
The spectrum of an FM channel multiplex signal, as it is generated by a transmitter before the modulation, is shown in this diagram.
The spectrum after the demodulation through a receiver is shown in the lower diagram.
Spectral parts, such as the pilot tone at 19~kHz, or the LR Diff part around 38~kHz can be seen clearly, as presented in the theory in Figure~\ref{fig:fm_channel_baseband_freqs}.
The demodulated spectrum shows an artifact at 57~kHz, which does not exist in the original transmitter spectrum.
This is an artifact of the demodulation process, specifically the frequency-shift of the LR Diff part to baseband.
The artifact is very low in energy, because it is attenuated by a BP filter in the down-conversion process, and can thus be ignored for further processing, even though it can still be seen in the spectral analysis.

\includepicture [1.0] [0] {Matlab analysis in the frequency domain. Spectral parts, like the mono audio, the pilot tone, or the LR difference signal around the 38~kHz subcarrier can clearly be seen.} {matlab_analysis_freq_domain} {img/matlab/matlab_analysis_freq_domain}
%TODO: latex font for this diagram!

@@

This section describes the implementation in GNU Radio.
A transmitter, as well as a receiver are implemented using this software.
Both can be deployed on actual hardware, so a continuous, live transmitter or receiver is built in a real environment.

@@

GNU Radio provides a comprehensive set of DSP blocks, which are available to build various functionalities.
Multiple blocks are connected to a block diagram, that is very similar to the one in Figure~\ref{fig:bd_matlab_transmitter}.
An external device, the USRP B200mini from Ettus Research, is used to broadcast the RF signal over the air.\\

@@

Special care needs to be taken when transmitting signals over the air.
The frequency spectrum is strongly limited by the radio regulations and may be monitored by the authorities.
The transmitter setup needs to be adapted carefully to the respective local regulations, to avoid any legal issues.
Some of the regulations are mentioned in Section~\ref{sec:FrequencyBand}, but may differ in various countries, however.\\

@@

\includepicture [0.9] [0] [H] {Block diagram of the GNU Radio transmitter setup.} {bd_gnuradio_setup_transmitter} {img/draw.io/bd_gnuradio_setup_transmitter}
%TODO: transmitted waves in reverse order (getting larger as they distance..)

@@

The receiver is also implemented by using standard DSP blocks that are available in GNU Radio.
The implemented structure however does not represent a stereo receiver.
Instead, a mono receiver is implemented.
The antenna signal is FM-demodulated, decimated and then directly sent to the speakers as the output, without any further processing.
This simple receiver is intended as a proof-of-concept and a first example of the cooperation of the external hardware and the GNU Radio software.
Therefor, the mono receiver is sufficient, as the focus of the GNU Radio application is put on the transmitter side.\\

This receiver is tested with two different external devices, the previously mentioned USRP B200mini and any RTL-SDR device.
The devices are explained into more detail later in Section \ref{sec:deployment-on-hw:gnu-radio-devices}.
A block diagram of the receiver application is shown in Figure~\ref{fig:bd_gnuradio_setup_receiver}.

%TODO: b200mini --> B200mini (upper-case 'B')
\includepicture [0.9] [0] [H] {Block diagram of the GNU Radio receiver setup.} {bd_gnuradio_setup_receiver} {img/draw.io/bd_gnuradio_setup_receiver}

@@

The usage of a software like GNU Radio in combination with external hardware is an interesting tool for a use-case like this project.
An FM receiver is to be implemented in multiple different versions and all of them need to be verified somehow -- this is where the SDR transmitter, created with GNU Radio and external hardware, opens up an entire new possibility from the viewpoint of system level design integration.\\

Previously it was explained, that it is not possible to use recorded antenna data for verification.
The reason was, that the underlying original data is not known and therefor a comparison of the original, versus the received, decoded data is not feasible.
Consequently, the Matlab transmitter model was justified by the fact, that its originally encoded audio data is well-known.\\

However, with the current approach of using a GNU Radio SDR transmitter, the originally encoded data of the received signal \textit{is} well-known, because the transmitted signal is generated locally.
Thus, the SDR can be used to transmit a verification signal.
This signal can be recorded to file by another SDR, or sourced into the DUT hardware directly.
